---
title: "Regression_Tree_Modeling"
author: "Alexander Thiersch"
date: "4/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Import libraries
library(ezids)
library(dplyr)
library(data.table)
library(readr)
library(lattice)
library(corrplot)
library(psych)
library(kableExtra)
library(ggpubr)
library(ggplot2)

#Regression Tree Packages
library(rpart)
library(rattle)
library(pROC)
library(tree)
library(MASS)
library(caTools)
library(ModelMetrics)
library(modelr)
library(randomForest)


```

# Technical Analysis

## Instructions

Technical analysis (R code), 25% (First attempt due: Mar 28 Monday 12:30pm)

Just before your team presentation, provide an R-markdown, or R script file, which shows the R-code and brief explanations as well as the rationale of the Exploratory Data Analysis of
your project. (Also submit your data file, or give the online source url.) This submission,
together with your final document (in part VI), will determine your team’s grade. This
document shows a technical person the math/stat/codes that you used in your analysis.

The document should include:

- Summary of the dataset
- Descriptive Statistics
- Graphical representations of the data
- [When applicable]  Measures of Variance / sd
- [When applicable]  Normality tests
- [When applicable]  Initial correlation / Chi Square tests / ANOVA analysis / Z-test or Z-interval / T-test or T-interval etc.

## Summary

The variables in the dataset are:

* `region`: name of the US Census Bureau region (name)
* `division`: name of the US Census Bureau division (contained with a census region)
* `state`: state abbreviation
* `statecode`: FIPS state code
* `countycode`: FIPS county code
* `fipscode`: 5-digit FIPS Code (county-level); combines `statecode` and `countycode`
* `county`: county name
* `year`: report release year from [County Health Rankings](https://www.countyhealthrankings.org/); range of 2016-2021
* `county_ranked`: Indicates whether or not the county was ranked; 0=unranked, 1=ranked, or NA for aggregated national or state-level data
* `mental_health_days`: Average number of mentally unhealthy days reported in past 30 days (age-adjusted)
* `mental_distress_rate`: Percentage of adults reporting 14 or more days of poor mental health per month
* `inequality`: Ratio of household income at the 80th percentile to income at the 20th percentile (Income inequality)
* `median_inc`: The income where half of households in a county earn more and half of households earn less
* `hs_grad`: Percentage of adults ages 25 and over with a high school diploma or equivalent
* `college`: Percentage of adults ages 25-44 with some post-secondary education
* `unempl`: Percentage of population ages 16 and older unemployed but seeking work
* `child_poverty`: Percentage of people under age 18 in poverty
* `single_parent`: Percentage of children that live in a household headed by single parent
* `severe_housing`: Percentage of households with severe housing problems
* `food_index`: Index of factors that contribute to a healthy food environment, from 0 (worst) to 10 (best)
* `mh_providers`: rate of providers to 100,000 population
* `pop_provider_ratio`: ratio of population to mental health providers (i.e., population served per provider)
* `pop`: census population estimate
* `pct_below18`: percent of population younger than 18
* `pct_black`: percent of population that are African-American or non-Hispanic Black
* `pct_native_am`: percent of population that are Native American or Alaska Natives
* `pct_asian`: percent of population that are Asian
* `pct_pacific`: percent of population that are Native Hawaiian or Other Pacific Islander
* `pct_hispanic`: percent of population that are Hispanic
* `pct_white`: percent of population that are non-Hispanic white or Caucasian
* `pct_female`: percent of population that are female
* `pct_rural`: percent of population that live in rural areas

For more information, see [the measures online](https://www.countyhealthrankings.org/explore-health-rankings/measures-data-sources/2021-measures).

```{r import, results='hide', echo=F}

# import csv as dataframe
dframe <- data.frame(read.csv("/Users/akthiersch/Desktop/Datasets/analytic_data_2016_2021_with_regions.csv"))
#View(dframe)

# convert variables to factor
dframe$region <- factor(dframe$region)
dframe$division <- factor(dframe$division)
dframe$statecode <- factor(dframe$statecode)
dframe$countycode <- factor(dframe$countycode)
dframe$fipscode <- factor(dframe$fipscode)
dframe$county_ranked <- factor(dframe$county_ranked)

# check structure of data
str(dframe)
#summary(dframe)

```

```{r summary, results='hide'}

# look at county_ranked var; not all counties are ranked; also some aggregated data per state and country exist in the observations
# =1 means they are ranked, =0 means unranked, and =NA is for state/national data
#print(summary(dframe$county_ranked))

# subset of dataframe including only ranked counties
ranked <- dframe %>% subset(county_ranked==1)

# subset of dataframe including only ranked counties
unranked <- dframe %>% subset(county_ranked==0)

# subset of dataframe including only aggregated data
aggregated <- dframe %>% subset(is.na(county_ranked))

# duplicate column and rename level labels for easier reading
ranked$region_abb <- ranked$region
levels(ranked$region_abb) <- c("", 
                              "MW",  # re-level factor labels
                              "NE",
                              "S", 
                              "W")

# subset ranked data by region
ranked_MW <- ranked %>% subset(region=="Midwest")
ranked_NE <- ranked %>% subset(region=="Northeast")
ranked_SO <- ranked %>% subset(region=="South")
ranked_WE <- ranked %>% subset(region=="West")

# subset ranked data into annual datasets
ranked16 <- ranked %>% subset(year==2016)
ranked17 <- ranked %>% subset(year==2017)
ranked18 <- ranked %>% subset(year==2018)
ranked19 <- ranked %>% subset(year==2019)
ranked20 <- ranked %>% subset(year==2020)
ranked21 <- ranked %>% subset(year==2021)

# view size of annual datasets
df_annual_list <- list(ranked16, ranked17, ranked18, ranked19, ranked20, ranked21)
for (df in df_annual_list) {
  print(paste("Observations in", median(df$year), "data: ", nrow(df)))
  
}

# sort dataframe
ranked <- ranked[order(ranked$year, ranked$region, ranked$division, ranked$statecode, ranked$countycode), ]

# view head and tail of ranked data
xkabledplyhead(ranked, 2, title = "Table: First 2 Rows of Ranked Data")
xkabledplytail(ranked, 2, title = "Table: Last 2 Rows of Ranked Data")

```

```{r}
#Well, one quick & dirty way could be to modify tree::partition.tree, which does that job in base graphics. Heres an example for an iris classification tree:

gg.partition.tree <- function (tree, label = "yval", ordvars, ...) 
{
    ptXlines <- function(x, v, xrange, xcoord = NULL, ycoord = NULL, 
        tvar, i = 1L) {
        if (v[i] == "<leaf>") {
            y1 <- (xrange[1L] + xrange[3L])/2
            y2 <- (xrange[2L] + xrange[4L])/2
            return(list(xcoord = xcoord, ycoord = c(ycoord, y1, 
                y2), i = i))
        }
        if (v[i] == tvar[1L]) {
            xcoord <- c(xcoord, x[i], xrange[2L], x[i], xrange[4L])
            xr <- xrange
            xr[3L] <- x[i]
            ll2 <- Recall(x, v, xr, xcoord, ycoord, tvar, i + 
                1L)
            xr <- xrange
            xr[1L] <- x[i]
            return(Recall(x, v, xr, ll2$xcoord, ll2$ycoord, tvar, 
                ll2$i + 1L))
        }
        else if (v[i] == tvar[2L]) {
            xcoord <- c(xcoord, xrange[1L], x[i], xrange[3L], 
                x[i])
            xr <- xrange
            xr[4L] <- x[i]
            ll2 <- Recall(x, v, xr, xcoord, ycoord, tvar, i + 
                1L)
            xr <- xrange
            xr[2L] <- x[i]
            return(Recall(x, v, xr, ll2$xcoord, ll2$ycoord, tvar, 
                ll2$i + 1L))
        }
        else stop("wrong variable numbers in tree.")
    }
    if (inherits(tree, "singlenode")) 
        stop("cannot plot singlenode tree")
    if (!inherits(tree, "tree")) 
        stop("not legitimate tree")
    frame <- tree$frame
    leaves <- frame$var == "<leaf>"
    var <- unique(as.character(frame$var[!leaves]))
    if (length(var) > 2L || length(var) < 1L) 
        stop("tree can only have one or two predictors")
    nlevels <- sapply(attr(tree, "xlevels"), length)
    if (any(nlevels[var] > 0L)) 
        stop("tree can only have continuous predictors")
    x <- rep(NA, length(leaves))
    x[!leaves] <- as.double(substring(frame$splits[!leaves, "cutleft"], 
        2L, 100L))
    m <- model.frame(tree)
    if (length(var) == 1L) {
        x <- sort(c(range(m[[var]]), x[!leaves]))
        if (is.null(attr(tree, "ylevels"))) 
            y <- frame$yval[leaves]
        else y <- frame$yprob[, 1L]
        y <- c(y, y[length(y)])
        if (add) 
            lines(x, y, type = "s", ...)
        else {
            a <- attributes(attr(m, "terms"))
            yvar <- as.character(a$variables[1 + a$response])
            xo <- m[[yvar]]
            if (is.factor(xo)) 
                ylim <- c(0, 1)
            else ylim <- range(xo)
            plot(x, y, ylab = yvar, xlab = var, type = "s", ylim = ylim, 
                xaxs = "i", ...)
        }
        invisible(list(x = x, y = y))
    }
    else {
        if (!missing(ordvars)) {
            ind <- match(var, ordvars)
            if (any(is.na(ind))) 
                stop("unmatched names in vars")
            var <- ordvars[sort(ind)]
        }
        lab <- frame$yval[leaves]
        if (is.null(frame$yprob)) 
            lab <- format(signif(lab, 3L))
        else if (match(label, attr(tree, "ylevels"), nomatch = 0L)) 
            lab <- format(signif(frame$yprob[leaves, label], 
                3L))
        rx <- range(m[[var[1L]]])
        rx <- rx + c(-0.025, 0.025) * diff(rx)
        rz <- range(m[[var[2L]]])
        rz <- rz + c(-0.025, 0.025) * diff(rz)
        xrange <- c(rx, rz)[c(1, 3, 2, 4)]
        xcoord <- NULL
        ycoord <- NULL
        xy <- ptXlines(x, frame$var, xrange, xcoord, ycoord, 
            var)
        xx <- matrix(xy$xcoord, nrow = 4L)
        yy <- matrix(xy$ycoord, nrow = 2L)
        return(
          list(
            annotate(geom="segment", x=xx[1L, ], y=xx[2L, ], xend=xx[3L, ], yend=xx[4L, ]),
            annotate(geom="text", x=yy[1L, ], y=yy[2L, ], label=as.character(lab), ...)
          )
        )
    }
}





```

```{r}
#Partition Graph
ggplot(ranked, 
       aes(inequality, mental_health_days, color=region)) + 
  geom_point() +
  gg.partition.tree(tree(mental_health_days ~ inequality + hs_grad, data=ranked), 
                    label="Region", color = "cyan")



```


```{r}
#View(ranked)
#str(ranked)

#Creating Train and Test Set - ranked
set.seed(123)
train = sample(1:nrow(ranked), nrow(ranked)/1.25)

#First Model for Number of Mentally Unhealthy Days (ranked; all vars.)
model_ment_helth_day_tree <- rpart(mental_health_days ~ inequality + median_inc + hs_grad + college + unempl + child_poverty + single_parent + severe_housing + food_index + mh_providers + pop_provider_ratio, ranked, subset=train, control = list(maxdepth=30, cp=0.002))

#Summary for Number of Mentally Unhealthy Days
summ_table1<-summary(model_ment_helth_day_tree)
summ_table1$variable.importance

#Plot Regression Tree
#rpart.plot(model_ment_helth_day_tree)
fancyRpartPlot(model_ment_helth_day_tree, main = "Model 1 - Target y: mental_health_days (all vars.)")


#Print Complexity Penalty
printcp(model_ment_helth_day_tree)

rsq.rpart(model_ment_helth_day_tree)

#Plotting Complexity Parameters
plotcp(model_ment_helth_day_tree, minline = TRUE, lty = 3, col = 1)

```



```{r}
#Prune Tree

#Prune Tree Model
#prune_model_ment_helth_day_tree=prune.rpart(model_ment_helth_day_tree, cp=0.209, best=5)

prune_model_ment_helth_day_tree=prune.rpart(model_ment_helth_day_tree, cp=model_ment_helth_day_tree$cptable[which.min(model_ment_helth_day_tree$cptable[,"xerror"]),"CP"], best=5)


#Summary of Pruned Tree
summ_table2<-summary(prune_model_ment_helth_day_tree)
summ_table2$variable.importance
#Print Complexity Penalty of Prune Tree
printcp(prune_model_ment_helth_day_tree)

#Plot Complexity Penalty of Prune Tree 
plotcp(prune_model_ment_helth_day_tree,minline = TRUE, lty = 3, col = 1)

rsq.rpart(prune_model_ment_helth_day_tree)

#Plot Pruned Tree
fancyRpartPlot(prune_model_ment_helth_day_tree)

```

```{r}
#Predicting and MSE for Regression Trees

#Prediction/Testing - Original Model
yhat=predict(model_ment_helth_day_tree,newdata=ranked[-train,])
test=ranked[-train,"mental_health_days"]

plot(yhat,test, main="Regression Tree: Test vs. Predicted Values", xlab="Predicted Values", ylab="Test Values")
abline(0,1)

#MSE
mean((yhat-test)^2)

#RMSE
sqrt(mean((yhat-test)^2))


#Prediction/Testing - Pruned Model
yhat_prune=predict(prune_model_ment_helth_day_tree,newdata=ranked[-train,])
test=ranked[-train,"mental_health_days"]

plot(yhat_prune,test, main="Pruned Regression Tree: Test vs. Predicted Values", xlab="Predicted Values", ylab="Test Values")
abline(0,1)

#MSE
mean((yhat_prune-test)^2)

#RMSE
sqrt(mean((yhat_prune-test)^2))


```


```{r}
#Random Forest

library(randomForest)

#Generate Random Forest Model - Num. of Mentally Unhealthy Days (all vars.)
set.seed(123)
bag_ment_hlth_days=randomForest(mental_health_days ~ inequality + median_inc + hs_grad + college + unempl + child_poverty + single_parent + severe_housing + food_index + mh_providers + pop_provider_ratio + region, data=ranked, subset=train, mtry=11, importance=TRUE)

bag_ment_hlth_days

#Predict/Test Random Forest Model
yhat.bag = predict(bag_ment_hlth_days,newdata=ranked[-train,])
plot(yhat.bag, test, main="Random Forest: Test vs. Predicted Values", xlab="Predicted Values", ylab="Test Values")
abline(0,1)

#MSE
mean((yhat.bag-test)^2)

#RMSE
sqrt(mean((yhat.bag-test)^2))


#Determine Variable Importance
importance(bag_ment_hlth_days)

#Plot Variable Importance 
varImpPlot(bag_ment_hlth_days, main = "Variable Importance Plots for Random Forest")

```

